# DOMC Difficulty Variance

## Initial Run
Here is a document showing the results of item families derived from a single DOMC stem. Essentially we treat each possible combination of options as a differential question just to see the amount of variance from a single DOMC stem.

In this first run we treat each different option combination as a different item, including those for people who never saw the correct response (making all these p-values 0)

```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(readr)
library(knitr)
library(ggplot2)
library(lemon)

setwd_thisdir <- function () {
  this.dir <- dirname(parent.frame(3)$ofile)
  setwd(this.dir)
}


hp_data = read_csv('data/domc_order_difficulty/Full Responses After Exclusions.csv')

hp_clean = hp_data %>% filter(item_type == 'domc', item_component_type == 'domc_option') %>% group_by(delivery_id, item_id) %>% mutate(option_order = paste0(option_presented, collapse = ""))

hp_summary = hp_clean %>% group_by(delivery_id, item_id) %>% summarize(item_total_seconds = max(item_total_seconds), item_order = max(option_order), score = max(score)) %>% group_by(delivery_id, item_id) %>% mutate(order = paste(sort(unlist(strsplit(as.character(item_order), ""))), collapse = "")) 

hp_items = hp_summary %>% group_by(item_id, order) %>% summarize(p_value = mean(score), count = n())

```

## Item 10B_v1

```{r, message=FALSE, warning=FALSE}
tenb_v1 = hp_items %>% filter(item_id == '10B_v1')
kable(tenb_v1)
```
### Histogram
```{r results='hide', message=FALSE, warning=FALSE}
theme_update(plot.title = element_text(hjust = 0.5))
ggplot(tenb_v1, aes(x=p_value)) + geom_histogram(binwidth=.01) + labs(title = "Item10B_v1") 
```

## Item 14b_v1

```{r, message=FALSE, warning=FALSE}
fourteenb_v1 = hp_items %>% filter(item_id == '14B_v1')
kable(fourteenb_v1, format = "markdown")
```
### Histogram

```{r results='hide', message=FALSE, warning=FALSE}
ggplot(fourteenb_v1, aes(x=p_value)) + geom_histogram(binwidth=.01) + labs(title = "Item14B_v1")+ xlim(-.01,1) + ylim(0,20)
```

## Remove correct response

We notice quickly that it is pretty difficult to interpret the p-values given that there is a significant amount of peole who never see the correct answer so the p-value for those "children" is 0. Here is what happens when we redefine what it means to be a different item. Instead of looking at every combination of seen options we are only going to look at distractors. This means that someone who saw the correct option and the first option (coded 01) will have "seen" the "same" question as someone who just saw the first distractor(coded 1).

It cleans things up a bit but it is important to recognize what is going on.



```{r results='hide', message=FALSE, warning=FALSE}

hp_summary$order_no_cr = gsub('0','',hp_summary$order)
hp_summary$order_no_cr = ifelse(hp_summary$order_no_cr == '', '0', hp_summary$order_no_cr)
hp_items_no_cr = hp_summary %>% group_by(item_id, order_no_cr) %>% summarize(p_value = mean(score), count = n())

```

## Item 10B_v1

```{r, message=FALSE, warning=FALSE}
tenb_v1 = hp_items_no_cr %>% filter(item_id == '10B_v1')
kable(tenb_v1)
```

### Histogram

```{r, message=FALSE, warning=FALSE}
theme_update(plot.title = element_text(hjust = 0.5))
ggplot(tenb_v1, aes(x=p_value)) + geom_histogram(binwidth=.01) + labs(title = "Item10B_v1") + xlim(-.01,1)
```

## Item 14b_v1

```{r, message=FALSE, warning=FALSE}
fourteenb_v1 = hp_items_no_cr %>% filter(item_id == '14B_v1')
kable(fourteenb_v1)
```

### Histogram

```{r results='hide', message=FALSE, warning=FALSE}
ggplot(fourteenb_v1, aes(x=p_value)) + geom_histogram(binwidth=.01) + labs(title = "Item14b_v1") + xlim(-.01,1)
```

## All Item Plots

```{r, fig.height = 15, fig.width = 7, results='hide', message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=20) 

ggplot(hp_items_no_cr, aes(x=p_value)) + geom_histogram(binwidth=.01) + labs(title = "All Items") + xlim(-.01,1) + facet_wrap(~item_id, ncol=3)


```



