# DOMC Difficulty Factors

The purpose of this chapter is to briefly address portions of the item format that may lead to systematic differences in difficulty for DOMC items

```{r}

library(dplyr)
library(readr)
library(knitr)
library(ggplot2)
library(lemon)

setwd_thisdir <- function () {
  this.dir <- dirname(parent.frame(3)$ofile)
  setwd(this.dir)
}


hp_data = read_csv('data/domc_order_difficulty/Full Responses After Exclusions.csv')

hp_clean = hp_data %>% filter(item_type == 'domc' | item_type == 'multiple_choice', item_component_type == 'domc_option' | item_component_type == 'final', !grepl('Survey', hp_data$item_id)) %>% group_by(delivery_id, item_id) %>% mutate(option_order = paste0(option_presented, collapse = ""))

hp_summary = hp_clean %>% group_by(delivery_id, item_id, item_type) %>% summarize(item_total_seconds = max(item_total_seconds), item_order = max(option_order), score = max(score)) %>% group_by(delivery_id, item_id) %>% mutate(order = paste(sort(unlist(strsplit(as.character(item_order), ""))), collapse = "")) %>% ungroup()

total_score = hp_summary %>% group_by(delivery_id) %>% summarize(total_score = sum(score))

hp_final = left_join(hp_summary, total_score)
hp_final$char_count = nchar(hp_final$order)

```

# Total Test Score and Options Seen

Simply running a logistic regression predicting if a person will get an item correct based on teh total number of options seen seems disingenous for a DOMC item. The reason being is that more proficient examinees will on average see more options as they are more likely to reject incorrect options. 


Here we look at some regression results

```{r}
mylogit <- glm(score ~ total_score + factor(item_type) + char_count, data = hp_final, family = "binomial")
summary(mylogit) 

mylogit <- glm(score ~ factor(item_type) + char_count + factor(item_type)*char_count, data = hp_final, family = "binomial")
summary(mylogit) 
```

