# Recall VS Deduction

In this chapter we will look at some differences between recall and deduction items on a Harry Potter assessment.

```{r cars}
library(dplyr)
library(readr)
library(knitr)
library(ggplot2)
library(lemon)
library(stringr)
setwd_thisdir <- function () {
  this.dir <- dirname(parent.frame(3)$ofile)
  setwd(this.dir)
}

hp_data = read_csv('data/domc_order_difficulty/Full Responses After Exclusions.csv')

hp_clean = hp_data %>% filter(item_type == 'domc' | item_type == 'multiple_choice', item_component_type == 'domc_option' | item_component_type == 'final', !grepl('Survey', hp_data$item_id)) %>% group_by(delivery_id, item_id) %>% mutate(option_order = paste0(option_presented, collapse = ""))

recall <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
deduction = c("13", "14", "15", "16", "17", "18", "19", "20", "21", "22")
hp_clean$deduction <- ifelse(grepl(paste(deduction,collapse="|"), hp_clean$item_id),1,0)

hp_items = hp_clean %>% group_by(item_id, item_type) %>% summarize(p_value = mean(score), count = n(), deduction = mean(deduction,na.rm=TRUE))
hp_items$item_number = str_extract(hp_items$item_id, "[0-9]+")
```

## Some Graphs

Here is the difference in p-value between DOMC and multiple choice for the same items.

```{r}

ggplot(hp_items, aes(x=item_type, y=p_value, fill=item_type)) + geom_bar(stat="identity")  + facet_wrap(~item_number)

```

Here is the difference in p-value between DOMC and multiple choice for the same items.

```{r}
just_mc = hp_items %>% filter(item_type == 'multiple_choice')
just_domc = hp_items %>% filter(item_type != 'multiple_choice')

differences = hp_items %>% group_by(item_type) %>% summarize(average_p_value = mean(p_value, na.rm=TRUE))
kable(differences)

```

The total difference in p-values between item_types is `r mean(just_mc$p_value) - mean(just_domc$p_value)`.

Multiple Choice items appear to be slightly more difficult.